### 2014-11-16 appendices/appendix-vc.tex

\subsection{ Expressions for 
\texorpdfstring{${\cal V}^{a,\ell}_{nm}(\bfk)$, 
${\cal C}^{\ell}_{nm}(\bfk)$ 
and
$({\cal C}^{\ell}_{nm}(\bfk))_{;\bfk}$
}{Vnm and Cnm}}\label{calpcalc}

Expanding the wave function in plane waves we obtain
\begin{align}\label{eni.1}
\psi_{n\bfk}(\bfr)=\sum_\bfG A_{n\bfk}(\bfG)e^{i(\bfk+\bfG)\cdot\bfr}
,
\end{align}
where $\{\bfG\}$ are the reciprocal basis vectors satisfying
$e^{\bfR\cdot\bfG}=1$, with $\{\bfR\}$ the translation vectors in real
space, and $A_{n\bfk}(\bfG)$ are the expansion coefficients. Using
$m_e\bfv=-i\hbar\bfgnabla$ into the equivalent of Eq.~\eqref{intj}
 we obtain,\cite{mendozaPRB06}
\begin{align}\label{eni.2}
\calbv^\ell_{nm}(\bfk)=
\frac{\hbar}{2m_e}
\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
(2\bfk+\bfG+\bfG')
\gd_{\bfG_\parallel \bfG'_\parallel}  
f^\ell(G_\perp-G'_\perp)
,
\end{align}   
where
\begin{align}\label{vnl.9}
f^\ell(g)=\frac{1}{L}\int_{z_\ell-\gD^b_\ell}^{z_\ell+\gD^f_\ell} e^{igz}dz  
 ,
\end{align}
with $f^{\ell*}(g)=f^\ell(-g)$. 
the reciprocal lattice vectors $\bfG$ are decomposed into components
parallel to the surface $\bfG_\parallel$, and perpendicular to the
surface $G_\perp \hat z$, so
that $\bfG = \bfG_\parallel + G_\perp\hat z$.
The double summation over the $\bfG$ vectors can be efficiently done by  
creating a pointer array to identify all the plane-wave coefficients  
associated with the same $G_\parallel$.  
 Likewise we obtain that
\begin{align}\label{eni.4}
\calc^\ell_{nm}(\bfk)=
\sum_{\bfG,\bfG'} A^*_{n\bfk}(\bfG')  A_{m\bfk}(\bfG)
\gd_{\bfG_\parallel \bfG'_\parallel} 
f_\ell(G_\perp-G'_\perp)
.
\end{align}  
If $\calc^\ell(z)=1$, $f_\ell(g)=\gd_{g0}$, and from Eqs.~\eqref{eni.2} and \eqref{eni.4}
we obtain the full-slab or bulk values
$\bfv_{nm}(\bfk)$ and
$\calc^\ell_{nm}(\bfk)=\gd_{nm}$.

Since for any function $F(\bfr)$, $[\bfr,F(\bfr)]=0$, using 
Eqs.~\eqref{rnminn}, \eqref{conmri3n}, and \eqref{gendevnn}, we obtain
\begin{align} 
(\calc^\ell_{nm})_{;\bfk}
&=
i
\sum_{q} 
 \left(\bfr_{nq}
\calc^\ell_{qm}
-
\calc^\ell_{nq}
\bfr_{qm}
\right) 
+i\bfr_{nm}(\calc^\ell_{mm}-\calc^\ell_{nn}) 
,
\end{align} 
where we remind the reader that $\bfr_{nm}$ are calculated through
Eq.~\eqref{chon.10} for LDA.