### 2014-12-04 sent/submission_letter.tex
sent to prb

In this article we formulate a theoretical approach of surface
second-harmonic generation from semiconductor surfaces based on the
length gauge and the one electron density operator.  We calculate the
nonlinear second-order surface susceptibility tensor within the
independent particle approximation. We include for the first time in one
unique formulation: (i) the scissors correction, needed to have the
correct value of the energy band gap, (ii) the contribution of the
nonlocal part of the pseudopotentials, routinely used in ab initio
band structure calculations, and (iii) the derivation for the
inclusion of the cut function, used to extract the surface response.
We test our scheme with a clean Si(001)2x1 reconstructed surface with
excellent results.  Therefore, the inclusion of the three
aforementioned contributions is very important and makes our scheme
unprecedented and opens the possibility to study surface
second-harmonic generation with more versatility while providing more
accurate results.  Most of the cited works appear in the Physical
Review Letters and Physical Review B, for which we think that this
article would have good visibility in Physical Review B as well.