### 2015-01-08 reply/Scissors.tex

Introduction :

However this can be a very expensive calculation and thus one uses the much simpler scissors operator scheme. 34–36
It allows to “open” the DFT-LDA gap to its correct experimental or GW value for most bulk semiconductors. 
This approximation has already been used in linear optical calculations for surfaces, see [PRL 76 2810 (1996)], thus improving the agreement with experimental results.

Results B After Refs. 64,65
This is justified by the fact that the surface states of the clean Si(001) surface investigated are found, 
from GW calculations[PRB 52 1905 (1995)], to shift rigidly and to keep their dispersion with respect to LDA.